---
title: "Lab 1 Activity"
author: PSYC 7804 - Spring 2026
format: pdf
editor: source
header-includes:
    - \usepackage{setspace}
---

\onehalfspacing


```{r include=FALSE}
library(tidyverse)
```


For this lab activity we will use the `attitude` dataset, which is automatically loaded in `R`. 

**Description:** A survey of the clerical employees of a large financial organization. The data are aggregated from the questionnaires of the approximately 35 employees for each of 30 (randomly selected) departments. The numbers give the percent proportion of favourable responses to seven questions in each department.



```{r}
#| echo: false
#| eval: true
#| warning: false

library(tidyverse)

info <- tibble(Variable = names(attitude),
               Description = c("Overall rating", 
                               "Handling of employee complaints",
                               "Does not allow special privileges", 
                               "Opportunity to learn", 
                               "Raises based on performance",
                               "Too critical", 
                               "Advancement"))
library(kableExtra)

kable(info) %>% 
  column_spec(column = 2, width = "5in") %>% 
  column_spec(column = 1, bold = T) %>% 
  row_spec(0, bold = T)
```


We will be looking at how positions that allow opportunities for learning (`learning`) predict overall rating of their department (`rating`). The data is already loaded into `R` as `attitude`.


**1.** Calculate means and standard deviations of the two variables.

```{r}
#| echo: false
#| eval: false

mean(attitude$learning)
mean(attitude$rating)

sd(attitude$privileges)
sd(attitude$rating)
```

**2.** create plots to describe the distribution of the two variables. Use a different plot for each variable.

```{r}
#| echo: false
#| eval: false

set_theme(theme_classic())

# strange distribution

ggplot(attitude,
       aes(x = learning)) +
       geom_density()

# not very normally distributed

ggplot(attitude,  
 aes(x = rating)) +
 geom_histogram(bins = 15,
                color = "black",
                linewidth = .8,
                fill = "#3492eb")

# strange distributions to be expected given the small sample size                

```

**3.** Create a scatterplot with `rating` on the $y$-axis and `learning` on the $x$-axis and draw a regression line. What sign do you expect the slope to have? 

```{r}
#| echo: false
#| eval: false

ggplot(attitude,  
 aes(x = learning, y = rating)) +
 geom_point() +
 geom_smooth(method = "lm", 
             se = FALSE)

# Positive sign           
```

**4.** Run a linear regression and write the regression equation. How would interpret $\beta_1$, the slope of `learning`? 

```{r}
#| echo: false
#| eval: false

reg <- lm(rating ~ learning, attitude)

summary(reg)

# rating = 28.1741 + 0.6468*learning

# the slope of learning means that for each unit increase in learning, we expect a 0.6468 units increase in rating of the department.
```

- What is the predicted value of `rating` when `learning` = 58?

```{r}
#| echo: false
#| eval: false

reg$coef[1] + reg$coef[2]*58

# 65.68981 
```


**5.** Create a residuals plot and evaluate whether the residuals are reasonably evenly distributed around the regression line.

```{r}
#| echo: false
#| eval: false

ggplot(attitude, 
  aes(x = learning, y = reg$resid)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  geom_smooth(method = "loess",
              se = FALSE)

# seems alright given that the sample size is a bit small              
```


**6.** Can you identify a datapoint that is "unexpected" according to the regression? Print the row of the dataset associated with this datapoint. Given the regression, what is "unusual" about this datapoint?

```{r}
#| echo: false
#| eval: false

# the residual with the highest absolute value is the most "unexpected"
(point <- which.max(abs(reg$resid)))


# this point has high learning opportunities but low overall rating. This is unlike the general linear trend described by the regression

attitude[point, c("rating", "learning")]
```





