---
title: "Lab 4 Activity"
author: PSYC 7804 - Spring 2026
format: pdf
editor: source
header-includes:
    - \usepackage{setspace}
---

\onehalfspacing


We will be looking at the `Chirot` data from the `car` package again. As a reminder, this is data on the [the 1907 Romanian peasant revolt](https://en.wikipedia.org/wiki/1907_Romanian_peasants%27_revolt){target="_blank"}. Find a descriptions of the variables in the table below:


```{r}
#| echo: false
#| eval: true
#| warning: false

library(car)
library(tidyverse)

info <- tibble(Variable = names(Chirot),
               Description = c("Intensity of the rebellion",
                               "Commercialization of agriculture",
                               "Traditionalism",
                               "Strength of middle peasantry",
                               "Inequality of land tenure"))

library(kableExtra)

kable(info) %>% 
  column_spec(column = 2, width = "300px") %>% 
  row_spec(0, bold = T) %>% 
  column_spec(1, bold = T)
```


run the following code to name the data you will be using as `dat`:

```{r}
#| echo: true
#| eval: true
#| warning: false

library(car)
dat <- Chirot
```


**1.** Treat `intensity` as the outcome and `tradition` and `inequality` as predictors and run a multiple regression. What is the $R^2$ of this model? What about the adjusted $R^2$? do you think that they are appreciably different?

```{r}
#| echo: false
#| eval: false

reg <- lm(intensity ~ tradition + inequality, dat)
summary(reg)$r.squared
summary(reg)$adj.r.squared
```



```{r}
#| echo: false
#| eval: false

# yes, they are reasonably different, around a 5% difference

```

**2.** The formula for adjusted $R^2$ is
$$R^2_{\text{adj}} = 1 - (1 - R^2) \frac{n - 1}{n - k - 1},$$

where $R^2$ is the original $R^2$ value, $n$ is the sample size, and $k$ is the number of predictors. In `R`, first define $R^2$, $n$, and $k$ as variables; then, calculate the $R^2_{\text{adj}}$ using the equation above.

```{r}
#| echo: false
#| eval: false

R_sq <- summary(reg)$r.squared
n <- nrow(dat)
k <- 2

1 - ((1 - R_sq))*(n - 1)/(n - k - 1)
```

- assuming all other variables stay the same, what happens to $R^2_{\text{adj}}$ if $n = 100$ and if $n = 1,000$? Calculate $R^2_{\text{adj}}$ assuming $n = 100$ and $n = 1,000$ and compare the results with the original unadjusted $R^2$.

```{r}
#| echo: false
#| eval: false

n <- 100

1 - ((1 - R_sq))*(n - 1)/(n - k - 1)

n <-  1000

1 - ((1 - R_sq))*(n - 1)/(n - k - 1)

# as n increases the adjusted Rsquare becomes closer and closer to the unadjasted
```


**3.** Compute the bootstrapped confidence intervals for the two slopes of the regression in question 1. Run 3 different models with 100, 500, and 2000 bootstraps (remember to set a seed!). For each of them, print the confidence intervals and create histograms of the 

```{r}
#| echo: false
#| eval: false

set.seed(7531)

boot_100 <- car::Boot(reg, R = 100)

confint(boot_model)
```


```{r}
#| echo: false
#| eval: false

hist(boot_100, 
     parm = c("tradition", "inequality"))
```



```{r}
#| echo: false
#| eval: false

set.seed(7531)

boot_500 <- car::Boot(reg, R = 500)

```


```{r}
#| echo: false
#| eval: false

confint(boot_500)
hist(boot_500, 
     parm = c("tradition", "inequality"))
```



```{r}
#| echo: false
#| eval: false

set.seed(7531)

boot_2000 <- car::Boot(reg, R = 2000)
```


```{r}
#| echo: false
#| eval: false

set.seed(7531)

confint(boot_2000)
hist(boot_2000, 
     parm = c("tradition", "inequality"))
```




- What differences do you see in the distributions of the bootstrapped slopes? Based on the confidence intervals, would you make any different conclusions about the significance of the slopes (the answer may vary depedning on the chosen seed)?

```{r}
#| echo: false
#| eval: false

# with my current seed, the 95%CI for the slope of inequuality does include 0. the Distribution of the slopes are also a bit strange when we only draw 100 bootstrap samples.

```