---
title: "Lab 11: Logistic Regression"
author: "Fabio Setti"
institute: "PSYC 6802 - Introduction to Psychology Statistics"
bibliography: Additional files/references.bib
csl: Additional files/apa.csl
title-slide-attributes:
  data-transition: "zoom"
  data-visibility: "uncounted"
format:
   revealjs:
      footer: "PSYC 6802 - Lab 11: Logistic Regression"
      width: 1280
      height: 720
      chalkboard: true
      slide-number: c/t 
      theme: Fabio_theme/Fabio_theme.scss
      navigation-mode: linear
      controls: false
      auto-stretch: false
      header-includes:
        - <script src="Fabio_theme/Fabio_theme.js"></script>

editor: source
build: local
---

## Today's Packages and Data 🤗 

:::: {.columns}
::: {.column width = '50%'}

```{r}
#| code-fold: true
#| eval: false
#| echo: true
#| code-line-numbers: false
#| code-summary: "Install Packages Code"
#| classes: code-125

install.packages("caret")
```

```{r}
#| eval: true
#| echo: true
#| code-line-numbers: false
#| warning: false
#| classes: code-150

library(caret)
library(tidyverse)
# I am setting a theme for all the plots
theme_set(theme_classic(base_size = 16, accent = "#7a0b80",
                        base_family = 'serif'))

```



::: {.panel-tabset}

### `ppcor`

The `caret` package [@Kuhn_etal_2024] is a package that focuses on streamlining the process of creating predictive models. It focuses on machine learning applications and we will use it to help us with logistic regression.

:::

:::

::: {.column width = '50%'}
<center style="padding-bottom: 21px;"> [Data]{.data-title} </center>

Today we will look at the `Titanic dataset`. This dataset contains information about the passengers on the [Titanic](https://en.wikipedia.org/wiki/Titanic){target='_blank'}. 
 
```{r}
#| eval: true
#| echo: true 
#| message: false
#| code-line-numbers: false
#| classes: code-125

dat <- rio::import("https://fabio-setti.netlify.app/data/Titanic-Dataset.csv")
```

```{r}
library(FabioFun)

labels <- c("Passenger ID",
            "Weather Survived or not: 0 = No, 1 = Yes",
            "Ticket class: 1 = 1st, 2 = 2nd, 3 = 3rd",
            "Name of the Passenger",
            "Gender",
            "Age in Years",
            "No. of siblings/spouses aboard the Titanic",
            "No. of parents/children aboard the Titanic",
            "Ticket number",
            "Passenger fare",
            "Cabin number",
            "Port of Embarkation: C = Cherbourg, Q = Queenstown, S = Southampton")

labels_react <- reactbl_col.desc(names = colnames(dat),
                                 desc =  labels,
                                 style = "cursor: pointer; text-decoration: none;",
                                 info_color = "#7a0b80")

reactable::reactable(dat,
                     defaultColDef = colDef(minWidth = 140),
                     style = list(fontFamily = "Work Sans, sans-serif", fontSize = ".8rem"),
                     pagination = FALSE, highlight = TRUE, height = 300, columns = labels_react)
```

:::
::::



## Predicting Survival: Linear Regression?

The `Survived` variable denoted whether a passenger survived the Titanic wreck. let's say that we think that whether someone survived or not may be related to how much they spent on their Titanic ticket, `Fare`. we *can* run linear regression and get some results:

:::: {.columns}
::: {.column width = '60%'}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

lin_reg <- lm(Survived ~ Fare, dat)
summary(lin_reg)

```


:::

::: {.column width = '40%'}

However, `Survived` is a binary variable, where only 2 outcome are possible (0 = *died*, 1 = *survived*). Here are the counts:

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

table(dat$Survived)
```


</br>


<center> You can probably intuit that a linear regression is not appropriate for a binary DV, but looking at a plot will make many of the reasons clear. </center>


:::
::::


## Plotting the regression



:::: {.columns}
::: {.column width = '60%'}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot Code"

ggplot(dat, aes(x = Fare, y = Survived)) +
  geom_point() +
  geom_smooth(method = "glm",  method.args = list(family = "binomial"), se = FALSE)
```


:::
::: {.column width = '40%'}


:::
::::




## References 

<div id="refs"> </div>