---
title: "Lab 3: Chi-Square"
author: "Fabio Setti"
institute: "PSYC 6802 - Introduction to Psychology Statistics"
bibliography: Additional files/references.bib
csl: Additional files/apa.csl
title-slide-attributes:
  data-transition: "zoom"
  data-visibility: "uncounted"
format:
   revealjs:
      footer: "Lab 3: Chi-Square"
      width: 1280
      height: 720
      chalkboard: true
      slide-number: c/t 
      theme: Fabio_theme/Fabio_theme.scss
      navigation-mode: linear
      controls: false
      auto-stretch: false
      header-includes:
        - <script src="Fabio_theme/Fabio_theme.js"></script>

editor: source
build: local
---


## Today's Packages and Data 🤗

:::: {.columns}
::: {.column width="50%"}

```{r}
#| code-fold: true
#| eval: false
#| echo: true
#| code-line-numbers: false
#| code-summary: "Install Packages Code"
#| classes: code-150


# run for packages that you have not installed yet
# install.packages("tidyverse")
install.packages("psych")
```

```{r}
#| eval: true
#| echo: true
#| code-line-numbers: false
#| warning: false
#| classes: code-150

library(psych)
library(tidyverse)
```

::: {.panel-tabset}
### `psych`

The `psych` package [@revelle2024a] includes many functions that help with common analyses that you would encounter in psychological research. This package will come up over and over in the Lab potion of this course. 

:::

::: {.column width="50%"}

<center style="padding-bottom: 21px;"> [Data]{.data-title} </center>

You can open the data without any downloads with the line below. see [Here](https://fabio-setti.netlify.app/psyc7804/slides%20files/lab%201#/how-i-will-fix-everyones-working-directory-issues){target="_blank"} for *why* and *how* I do it. 
 
```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

WH_2024 <- import("https://fabio-setti.netlify.app/data/World_happiness_2024.csv")
```

<div style="font-size: 22px"> We'll use the [2024 world happiness report](https://www.worldhappiness.report/ed/2024){target="_blank"} data again. </div>

```{r}
reactable::reactable(WH_2024,
                     style = list(fontFamily = "Work Sans, sans-serif", fontSize = "1rem"),
                     pagination = FALSE, highlight = TRUE, height = 300)
```

:::
::::



