---
title: "Lab 9: Multiple Regression I"
author: "Fabio Setti"
institute: "PSYC 6802 - Introduction to Psychology Statistics"
bibliography: Additional files/references.bib
csl: Additional files/apa.csl
title-slide-attributes:
  data-transition: "zoom"
  data-visibility: "uncounted"
format:
   revealjs:
      footer: "PSYC 6802 - Lab 9: Multiple Regression I"
      width: 1280
      height: 720
      chalkboard: true
      slide-number: c/t 
      theme: Fabio_theme/Fabio_theme.scss
      navigation-mode: linear
      controls: false
      auto-stretch: false
      header-includes:
        - <script src="Fabio_theme/Fabio_theme.js"></script>

editor: source
build: local
---


## Today's Packages and Data 🤗

:::: {.columns}
::: {.column width = '50%'}

```{r}
#| code-fold: true
#| eval: false
#| echo: true
#| code-line-numbers: false
#| code-summary: "Install Packages Code"
#| classes: code-125


install.packages("GGally")
install.packages("devtools")
devtools::install_github("quinix45/FabioFun")
```

```{r}
#| eval: true
#| echo: true
#| code-line-numbers: false
#| warning: false
#| classes: code-150

# package 'devtools' must be installed first
library(GGally)
library(tidyverse)
library(FabioFun)
# I am setting a theme for all the plots
theme_set(theme_classic(base_size = 16, accent = "#7a0b80",
                        base_family = 'serif'))
```



::: {.panel-tabset}

### `GGally`

The GGally package [@Schloerke_etal_2024] builds upon `ggplot2` and includes many functions for creating complex plots. We will be using it to create scatterplot matrices.

### `FabioFun`

As you may have guessed, this is my package 🫣 I use it both to save some functions that I find may be useful in the future. I'll show some functions to create interactive 3D plots and APA style regression tables. 

### `flextable`

The `flextable` package [@Gohel_etal_2024] allows to create highly customizable tables. The function that I use to create APA style Word tables makes heavy use of this package.

:::

:::

::: {.column width = '50%'}
<center style="padding-bottom: 21px;"> [Data]{.data-title} </center>

Today we'll be looking at data from the [2024 World Happiness report](https://www.worldhappiness.report/ed/2024/){target='_blank'}. 
 
```{r}
#| eval: true
#| echo: true 
#| message: false
#| code-line-numbers: false
#| classes: code-125

dat <- rio::import("https://fabio-setti.netlify.app/data/World_happiness_2024.csv")
```

```{r}
library(FabioFun)


labels <- c("Country name",
            "Region of the World",
            "Country happiness score. COrresponds to the sum of the other variables",
            "Log of country Gross Domestic Product (GDP)",
            "Perceived Social Support in the Country",
            "Country Healthy life Expectancy",
            "Perceived Freedom in the Country",
            "Perceived Generosity in the Country",
            "Perceived Corruption of the Country")

labels_react <- reactbl_col.desc(names = colnames(dat),
                                 desc =  labels,
                                 style = "cursor: pointer; text-decoration: none;",
                                 info_color = "#7a0b80")

reactable::reactable(dat,
                     defaultColDef = colDef(minWidth = 170),
                     style = list(fontFamily = "Work Sans, sans-serif", fontSize = "1rem"),
                     pagination = FALSE, highlight = TRUE, height = 300, columns = labels_react)
```

:::
::::


## Predicting Country Happiness

In our data we have many variables that are likely related to happiness. I will just pick 2 of them for the purpose of this Lab. Let's say that we want to evaluate how `Freedom` and `Corruption` predict `Happiness`. If we run two separate regressions, we find that both have reasonably high $R^2$ values:

:::: {.columns}
::: {.column width = '50%'}

::: {.fragment fragment-index=1}


```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-115

reg_free <- lm(Happiness_score ~ Freedom, dat)
summary(reg_free)
```

:::

:::
::: {.column width = '50%'}

::: {.fragment fragment-index=2}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-115

reg_corr <- lm(Happiness_score ~ Corruption, dat)
summary(reg_corr)
```

:::

:::
::::


## Well, That's just the correlation 🤷

As we have learned, regression with one predictor is just a contrived way of calculating correlations among variables. We can better visualize the relation among out 3 variables with the `ggpairs()` function from `GGally`:

::: {.fragment fragment-index=1}

:::: {.columns}
::: {.column width = '35%'}

```{r}
#| eval: false
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

# here I am selecting the 3 columns that I wantr to plot
ggpairs(dat[,c(3, 7,9)])
```

- **Lower part:** scatterplot between variables
- **Diagonal:** variable distribution
- **Upper part:** correlation between variables

Notably, `Freedom` and `Corruption` are also correlated.

:::
::: {.column width = '65%'}

```{r}
#| eval: true
#| echo: false
#| code-line-numbers: false
#| classes: code-125

ggpairs(dat[,c(3, 7,9)])
```

:::
::::

:::

::: {.fragment fragment-index=2}

Separate regressions miss the points that our predictors share information!

:::

## Multiple Independent Variables

:::: {.columns}
::: {.column width = '52%'}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

reg_multi <- lm(Happiness_score ~ Freedom + Corruption, dat)
summary(reg_multi)
```

:::
::: {.column width = '48%'}

We can include both predictors in our regression by adding variables in the `lm()` function like so `lm(DV ~ IV1 + IV2 + ..., dat)`. for now I will note two things:


::: {.fragment fragment-index=1}
1. Both slopes have changed and are not as large as they were in the individual regressions. The regression equations is now:

 <p style='font-size: 20px'>   $$\mathrm{Happiness} = 5.09 + 4.03 \times \mathrm{Freedom} - 2.44 \times \mathrm{Corruption}$$ </p>

:::


::: {.fragment fragment-index=2}
2. The $R^2$ is roughly $.48$, which is lower than the sum of the $R^2$ from the two individual regressions from before ($.41 + .2 = .
61$).
:::

:::
::::

::: {.fragment fragment-index=3}

What is going on here? 🤔 a graphical illustration may provide some intuition!

:::

## Individual regression plots

Below I plot the individual regression lines for `Freedom` and `Corruption` predicting `Happiness`. 

:::: {.columns}
::: {.column width = '50%'}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot Code"

ggplot(dat,  
 aes(x = Freedom, y = Happiness_score)) +
 geom_point() + 
    geom_smooth(method = "lm",  
                se = FALSE)
```

:::
::: {.column width = '50%'}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot Code"

ggplot(dat,  
 aes(x = Corruption, y = Happiness_score)) +
 geom_point() + 
    geom_smooth(method = "lm",  
                se = FALSE)
```



:::
::::

::: {.fragment fragment-index=1}

Now, what happens if we try to overlay the regression lines that we got from our multiple regression? Will they "fit the data" better that the regression lines shown here?

:::



## Adding Multiple regression Lines

if we add our multiple regression lines to the plot... (🥁 *drum roll*)

::: {.fragment fragment-index=1}

:::: {.columns}
::: {.column width = '50%'}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot Code"

ggplot(dat,  
 aes(x = Freedom, y = Happiness_score)) +
 geom_point() + 
    geom_smooth(method = "lm",  
                se = FALSE) +
    geom_abline(intercept = coef(reg_multi)[1], 
                slope = coef(reg_multi)[2])

```

:::
::: {.column width = '50%'}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot Code"

ggplot(dat,  
 aes(x = Corruption, y = Happiness_score)) +
 geom_point() + 
    geom_smooth(method = "lm",  
                se = FALSE)+
    geom_abline(intercept = coef(reg_multi)[1], 
                slope = coef(reg_multi)[3])

```

:::
::::

:::

::: {.fragment fragment-index=2}

Wait, we are way off now 🤨 And here I thought multiple regression would be better 😑 ...Unsurprisingly, the multiple regression is fine, our *perspective* is the problem!

:::

## A better Perspective

We have *3* variables, `Freedom`, `Corruption`, and `Happiness`, but we are looking at them in *2* dimensions with the individual scatterplots. We need a **3D** scatterplot!



:::: {.columns}
::: {.column width = '40%'}

::: {.fragment fragment-index=1}

```{r}
#| eval: false
#| echo: true 
#| code-line-numbers: false
#| classes: code-125


# run "devtools::install_github("quinix45/FabioFun")" if you haven't installed the package yet
library(FabioFun)

nice_3D_plot(y = dat$Happiness_score,
             x1 = dat$Freedom,
             x2 = dat$Corruption,
             dot_labels = dat$Country_name,
             axis_names = c("Happiness", 
                            "Freedom", 
                            "Corruption"))
```

What is happening is that the scatterplots are *only the sides* of this box ‼️

:::

::: {.fragment fragment-index=2}
Multiple regression works on this plot on the right, not the individual ones; multiple regression *operates on a different dimension*.
:::

:::
::: {.column width = '60%'}

::: {.fragment fragment-index=1}

```{r out.width = "100%"}

library(FabioFun)


nice_3D_plot(y = dat$Happiness_score,
             x1 = dat$Freedom,
             x2 = dat$Corruption,
             dot_labels = dat$Country_name,
             axis_names = c("Happiness", 
                            "Freedom", 
                            "Corruption")) %>%  
                    bslib::card(full_screen = TRUE)
```

:::

:::
::::


## The Regression *Plane*

The results from multiple regression create a **regression plane**, which is the 3D version of a regression line. 


:::: {.columns}
::: {.column width = '40%'}

::: {.fragment fragment-index=1}

```{r}
#| eval: false
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

nice_3D_plot(y = dat$Happiness_score,
             x1 = dat$Freedom,
             x2 = dat$Corruption,
             dot_labels = dat$Country_name,
             axis_names = c("Happiness", 
                            "Freedom", 
                            "Corruption"),
             reg_plane = TRUE)
```


The plane works exactly as the regression line. its surface is the best predition of `happiness` given a value of `Freedom` and `Corruption`. The residuals are now the distance from any of the dots to the plane! 

:::


::: {.fragment fragment-index=2}
**Question:** What would the plot look like with one extra variable?
:::

:::

::: {.column width = '60%'}

::: {.fragment fragment-index=1}

```{r out.width = "100%"}

nice_3D_plot(y = dat$Happiness_score,
             x1 = dat$Freedom,
             x2 = dat$Corruption,
             dot_labels = dat$Country_name,
             axis_names = c("Happiness", 
                            "Freedom", 
                            "Corruption"),
             reg_plane = TRUE) %>%  
                    bslib::card(full_screen = TRUE)
```

:::

:::
::::


## Added Variable Plots

If the 3D plot is a bit much for you, you can also visualize the results of multiple regression with **added variable plots**. To do that, we can pass out `lm()` object to the `avPlots()` function from `car`:


:::: {.columns}
::: {.column width = '70%'}

::: {.fragment fragment-index=1}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

car::avPlots(reg_multi)
```

:::

:::
::: {.column width = '30%'}

::: {.fragment fragment-index=2}
Essentially, added variable plots give you a visual representation of the relation between an IV and the DV *after* accounting for all other IVs.
:::

::: {.fragment fragment-index=3}
Added variable plots have the advantage of being able to show multiple regression results when more than 2 IVs are involved. Why? Because with 3 IVs you would need a 4D plot, and humans can't really do that 😔 
:::

:::
::::

## Interpreting Multiple Regression

:::: {.columns}
::: {.column width = '50%'}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

summary(reg_multi)
```


:::
::: {.column width = '50%'}

Now that we have looked at our results graphically for a while, let's interpret the numbers that we get.

::: {.fragment fragment-index=1}

- **intercept**: the expected value of `Happiness` when both `Freedom` and `Corruption` are at 0.


- **Slope of `Freedom`**: the expected change in `Happiness` for 1-unit increase in `Freedom`, controlling for `Corruption`. 

- **Slope of `corruption`**: the expected change in `Happiness` for 1-unit increase in `Corruption`, controlling for `Freedom`.

- The $R^2$ is the variance in `Happiness` joiontly explained by `Corruption` and `Freedom`.

:::

:::
::::

::: {.fragment fragment-index=5}
**Note:** Instead of "controlling for", you may also hear "accounting for" or "holding constant". They are all the same thing. 
:::


## Checking Residuals

Because now we have multiple IVs, we need need to check the residuals against each of them. We can use the `residualPlots()` function from `car` to quickly get residual plots for all of our predictors:

:::: {.columns}
::: {.column width = '70%'}

::: {.fragment fragment-index=1}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

car::residualPlots(reg_multi, fitted = FALSE, test = FALSE)

```

:::

:::
::: {.column width = '30%'}

::: {.fragment fragment-index=2}

<p style='font-size: 32px'>  As far as residuals go, I would say that these two plots are not too bad.  </p>


<p style='font-size: 32px'>
The residuals of `Freedom` probably look better than those of `Corruption` mostly because `Corruption` has more points clustered at higher values and few points at lowe values.
</p>

:::

:::
::::

## The Scale of Our Variables and Interpretation

:::: {.columns}
::: {.column width = '50%'}



Our regression intercept and slope are:

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

coef(reg_multi)
```

::: {.fragment fragment-index=1}

If we look at the range of our variables, it should be apparent why the "1-unit increase in the IV" is not very meaningful in our case:

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

range(dat$Happiness_score)
range(dat$Freedom)
range(dat$Corruption)
```

:::

:::

::: {.column width = '50%'}

::: {.fragment fragment-index=2}

Because of the way `Freddom` and `Happiness` are measures, a "1-unit" increase is already out of range, making the value of the slopes not very representative of the data.

Once again, standardizing our variables can help us produce more interpretable results:

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

dat$Happiness_std <- scale(dat$Happiness_score)[,1]
dat$Freedom_std <- scale(dat$Freedom)[,1]
dat$Corruption_std <- scale(dat$Corruption)[,1]

```

:::

::: {.fragment fragment-index=3}

</br>

And now all the variables are on the same measurement unit. As a reminder, a value of 0 now represents the mean of the variable, and a value of -1/+1 represents a value 1 SD below/above the mean.
:::

:::
::::

## Standardized Multiple Regression


:::: {.columns}
::: {.column width = '60%'}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

reg_std <- lm(Happiness_std ~ Freedom_std + Corruption_std, dat)
summary(reg_std)
```

:::
::: {.column width = '40%'}

Now that the variables have been standardized, when can interpret the slopes as follows:

- When controlling for `Corruption`, 1 SD increase in `Freedom` will result in a $0.55$ SD units *increase* in `Happiness`.

- When controlling for `Freedom`, 1 SD increase in `Corruption` will result in a $0.26$ SD units *decrease* in `Corruption`.

:::
::::

::: {.fragment fragment-index=1}

In general, I believe you should always report standardized regression results, because your readers always know the scale of all your variables and how to interpret regression coefficients. 

:::

## APA Style Table for Regression Results

because I hate creating tables manually, at some point I made function that does that for me ✨ Also, creating big tables by hand is generally not a good idea: you waste time and likely make typos.

:::: {.columns}
::: {.column width = '70%'}

::: {.fragment fragment-index=1}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

library(FabioFun)
reg_tab <- regression_flextable(reg_std, 
                                var_names = c("Freedom", "Corruption"), 
                                intercept = FALSE)
reg_tab
```

:::

::: {.fragment fragment-index=2}

```{r}
#| eval: false
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

# save to Word document (will be saved to current working directory)
flextable::save_as_docx(reg_tab, path = "table.docx")
```


Feel free to use this instead of creating regression tables manually.

:::

:::
::: {.column width = '30%'}

::: {.fragment fragment-index=1}
the `regression_flextable()` function I made takes in an `lm()` object and has some optional arguments (see help page). Here I rename the regression coefficients with the `var_names =` argument.

:::


::: {.fragment fragment-index=2}
You can save the table as a `.docx` file (Word document) by using the `save_as_docx()` function from the `flextable()` package. You specify the file name in the `path =` argument.
:::

:::
::::


## Categorical Variables 

So far, we have only seen categorical IVs in ANOVA. Because ANOVA is a regression, we can indeed use categorical predictors in a regression. The catch is that we need to trick the regression into thinking that the categorical predictors are numeric variables 🧐

:::: {.columns}
::: {.column width = '70%'}

::: {.fragment fragment-index=1}

The one categorical variable we have in our data is `Region`, which describes what world region the country is in. 

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

table(dat$Region)
```

:::

:::
::: {.column width = '30%'}

::: {.fragment fragment-index=2}

<p style='font-size: 30px'>  **Idea:** we think that if we use the mean of `Happiness` of each of these regions, we will improve our 
 predictions! (this should sound familiar) </p>

:::

:::
::::


::: {.fragment fragment-index=3}
Now, we need to somehow express the `Region` variable in some numeric fashion, such that the regression switches to using the mean of `Happiness` by `Region` when it's predicting `Happiness` of a country from that region 💡
:::

::: {.fragment fragment-index=4}
There actually are many ways of doing this, but we will make use of the fact that regression slopes are the "change in `Y` per 1-unit change in `X`". We will focus on the trick known as **dummy coding**!
:::

## Dummy Coding With Factor Variables


Factor variables have the advantage of automatically doing dummy coding for you! I will filter for just a few regions to make the example a bit easier to follow:

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

dat_filt <- dat %>% 
   filter(Region %in% c("Africa", "Asia", "Western Europe", "Eastern Europe"))

```


:::: {.columns}
::: {.column width = '50%'}


::: {.fragment fragment-index=1}

Let's turn `Region` into a factor variable and investigate first:

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

dat_filt$Region <- factor(dat_filt$Region)
```

:::

::: {.fragment fragment-index=2}

We can now use the `contrasts()` function to see that `R` has done dummy coding for us behind the scenes:

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

contrasts(dat_filt$Region)
```

:::

:::


::: {.column width = '50%'}

::: {.fragment fragment-index=3}

If we look at each row, you can see each variable is assigned a vector of $0$s and $1$s. Crucially, each vector uniquely identifies the region (e.g., `0, 0, 0` corresponds to `Africa`)

:::

::: {.fragment fragment-index=4}

If we look at the columns, we can see that a $1$ appears on when the column name corresponds to the row name. Also, there is no column for `Africa` because we don't need it, just leaving `0,0,0` will automatically represent `Africa`.

:::

::: {.fragment fragment-index=5}

What does this mean for a regression?

:::

:::
::::


## Regression With a Categorical Variable

:::: {.columns}
::: {.column width = '60%'}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

reg_cat <- lm(Happiness_score ~ Region, dat_filt)
summary(reg_cat)
```

:::
::: {.column width = '40%'}

::: {.fragment fragment-index=1}

We get 3 "slopes" each corresponding to the regions that did not have all 0s. Let's look at the mean `Happiness` for each region:

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

dat_filt %>% 
   group_by(Region) %>% 
    summarise(Happ_mean = mean(Happiness_score))

```

:::

:::
::::

::: {.fragment fragment-index=2}

You should see that the intercept is the mean `Happiness` for `Africa`, while the "slopes" are the differences from each of the other regions `happiness` from that of `Africa`! 

:::

## The Graphical intuition

What is happening is that we have a *reference group*, which is the one coded a all 0s, `Africa` in our case. All other slopes will be the mean differences from the reference group. 

:::: {.columns}
::: {.column width = '30%'}

::: {.fragment fragment-index=1}

the intercept and the slopes are:

```{r}
#| eval: true
#| echo: false 
#| code-line-numbers: false
#| classes: code-125




dat_coef <- data.frame(c("Intercept", "Asia", "Eastern Europe", "Western Europe"), 
                        as.numeric(round(coef(reg_cat), 2)))
colnames(dat_coef) <- c("Coefficient", "Value")
dat_coef
```


The slopes represent the change in mean `Happiness` for each region from the reference group! 

:::

:::
::: {.column width = '70%'}

::: {.fragment fragment-index=1}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot Code"

happ_means <- dat_filt %>% 
   group_by(Region) %>% 
    summarise(Happ_mean = mean(Happiness_score))

# reference mean 
ref <- as.numeric(happ_means[1,2])

ggplot(happ_means, aes(x = Region, y = Happ_mean)) +
  geom_point(size = 4, color = "#7a0b80") +
  geom_text(aes(label = round(Happ_mean, 2)), vjust = -1) +
  geom_hline(yintercept = ref, lty = 3) +
  geom_segment(x = "Asia", 
               xend = "Asia", 
               y = ref,
               yend = as.numeric(happ_means[2,2]), lty = 3) +
 geom_segment(x = "Eastern Europe", 
               xend = "Eastern Europe", 
               y = ref,
               yend = as.numeric(happ_means[3,2]), lty = 3) +
geom_segment(x = "Western Europe", 
               xend = "Western Europe", 
               y = ref,
               yend = as.numeric(happ_means[4,2]), lty = 3) +
annotate("text", x = 1,  y = 4.2, label = "Reference Point") +
annotate("text", x = 2.2,  y = 5.3, label = paste("+", round(coef(reg_cat)[2], 2))) +
annotate("text", x = 3.2,  y = 5.3, label = paste("+", round(coef(reg_cat)[3], 2))) +
annotate("text", x = 4.2,  y = 5.3, label = paste("+", round(coef(reg_cat)[4], 2))) +
ylim(c(4, 6.9)) +
ylab("Mean Region Happiness")

```

:::

:::
::::


## Why Does this work?

I thin it is worthwhile to dive a bit more into *why* dummy coding works. The regression from two slides ago implied the following regression equation:

$$\mathrm{Happiness} = 4.40 + 1.09 \times \mathrm{Asia} + 1.55 \times \mathrm{Eastern Europe} +  2.35 \times \mathrm{Western Europe}$$


:::: {.columns}
::: {.column width = '50%'}

::: {.fragment fragment-index=1}

if you recall, the intercept is "the expected value of `Y` when all other variables are 0". If we apply it to this equation we get $4.40$, the *mean happiness of `Africa`*. This is the case because `Africa` is represented by `0,0,0`. 

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

intercept <- coef(reg_cat)[1]
slopes <- coef(reg_cat)[-1]

# the intercept + all slopes multiplied by 0. Will give the mean happiness for Africa
intercept + sum(slopes*c(0,0,0))
```

:::

:::

::: {.column width = '50%'}

::: {.fragment fragment-index=2}

Similarly, we know that `Western Europe` is represented as `0, 0, 1`. So, to get the mean `Happiness` of `Western Europe`, we run:

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

# mean happiness of western Europe (equivalent to doing 4.40 + 2.35)
intercept + sum(slopes*c(0,0,1))
```

For "1-unit increase" in `Western Europe`, the mean of `Happiness` goes from the reference of $4.40$ to $6.75$.

:::

:::
::::

## The Moral of the Story


So, dummy coding is one way of tricking regression into treating categorical variables as continuous. <u> Dummy ccoding is a type of contrast </u>, and as we have seen in [Lab 7](https://fabio-setti.netlify.app/psyc6802/lab_7#/significance-of-contrast){target='_blank'}, the numbers that we choose for contrasts can be whatever we like as long as they proportionally represent the comparison we are interested in.

:::: {.columns}
::: {.column width = '50%'}

::: {.fragment fragment-index=1}

For example, nothing prevents me from changing the $1s$ in the dummy coding to $3s$. The significance tests will be equivalent:

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false

contrasts(dat_filt$Region)[contrasts(dat_filt$Region) == 1] <- 3
contrasts(dat_filt$Region)
```

:::

::: {.fragment fragment-index=3}

However, this is impractical! Our coefficients are now interpreted as $\frac{1}{3}$ the mean difference comepared to the reference group. No one would want that 🤷

:::

:::
::: {.column width = '50%'}

::: {.fragment fragment-index=2}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: height-125


reg_cat <- lm(Happiness_score ~ Region, dat_filt)
summary(reg_cat)
```

:::

:::
::::


## References 

<div id="refs"> </div>