---
title: "Lab 5: Power"
author: "Fabio Setti"
institute: "PSYC 6802 - Introduction to Psychology Statistics"
bibliography: Additional files/references.bib
csl: Additional files/apa.csl
title-slide-attributes:
  data-transition: "zoom"
  data-visibility: "uncounted"
format:
   revealjs:
      footer: "PSYC 6802 - Lab 5: Power"
      width: 1280
      height: 720
      chalkboard: true
      slide-number: c/t 
      theme: Fabio_theme/Fabio_theme.scss
      navigation-mode: linear
      controls: false
      auto-stretch: false
      header-includes:
        - <script src="Fabio_theme/Fabio_theme.js"></script>

editor: source
build: local
---

## Today's Packages and Data 🤗

:::: {.columns}
::: {.column width = '50%'}

```{r}
#| code-fold: true
#| eval: false
#| echo: true
#| code-line-numbers: false
#| code-summary: "Install Packages Code"
#| classes: code-125

# run for packages that you have not installed yet
# install.packages("tidyverse")

install.packages("pwr")
```

```{r}
#| eval: true
#| echo: true
#| code-line-numbers: false
#| warning: false
#| classes: code-150

library(tidyverse)
library(pwr)
# I am setting a theme for all the plots
theme_set(theme_classic(base_size = 16, 
                        base_family = 'serif'))
```

</br>

::: {.panel-tabset}

### `pwr`

The `pwr` package [@champely_etal_2020] includes functions to conduct power analysis for some statistical analyses and experimental designs. Click [here](https://cran.r-project.org/web/packages/pwr/vignettes/pwr-vignette.html){target='_blank'} for examples on all the types of power calculations that the package allows for.
:::

:::

::: {.column width = '50%'}

<center style="padding-bottom: 21px;"> [Data]{.data-title} </center>

We won't look at data too much today, but let's load the data from last lab, which will come up for *retrospective* power analysis.

```{r}
#| eval: true
#| echo: true 
#| message: false
#| code-line-numbers: false
#| classes: code-125

dat <- rio::import("https://fabio-setti.netlify.app/data/Cuisine_rating.csv")
```

```{r}
reactable::reactable(dat,
                     style = list(fontFamily = "Work Sans, sans-serif", fontSize = "1rem"),
                     pagination = FALSE, highlight = TRUE, height = 300)

```

:::
::::

## References 

<div id="refs"> </div>