---
title: "Lab 8: Factorial ANOVA"
author: "Fabio Setti"
institute: "PSYC 6802 - Introduction to Psychology Statistics"
bibliography: Additional files/references.bib
csl: Additional files/apa.csl
title-slide-attributes:
  data-transition: "zoom"
  data-visibility: "uncounted"
format:
   revealjs:
      footer: "PSYC 6802 - Lab 8: Factorial ANOVA"
      width: 1280
      height: 720
      chalkboard: true
      slide-number: c/t 
      theme: Fabio_theme/Fabio_theme.scss
      navigation-mode: linear
      controls: false
      auto-stretch: false
      header-includes:
        - <script src="Fabio_theme/Fabio_theme.js"></script>

editor: source
build: local
---


## Today's Packages and Data 🤗

:::: {.columns}
::: {.column width = '50%'}

```{r}
#| code-fold: true
#| eval: false
#| echo: true
#| code-line-numbers: false
#| code-summary: "Install Packages Code"
#| classes: code-125


# run for packages that you have not installed yet
install.packages("emmeans")
```

```{r}
#| eval: true
#| echo: true
#| code-line-numbers: false
#| warning: false
#| classes: code-150


library(tidyverse)
library(emmeans)
# I am setting a theme for all the plots
theme_set(theme_classic(base_size = 16, accent = "#7a0b80",
                        base_family = 'serif'))
```



::: {.panel-tabset}

### `emmeans`

The `emmeans` package [@Lenth_etal_2025] is a great package for exploring the results of many linear, generalized linear, and mixed models. it also helps compute mean contrasts, trends, and comparisons of slopes. I discover some new functionality every time I look at this package!

:::

:::

::: {.column width = '50%'}
<center style="padding-bottom: 21px;"> [Data]{.data-title} </center>


I got the data from [text](https://people.hsc.edu/faculty-staff/blins/classes/spring17/math222){target='_blank'}

Today we will work with some data from [NYC public schools](https://www.kaggle.com/datasets/nycopendata/high-schools/data){target='_blank'}. Namely, we will focus on the average SAT scores by school across the different boroughs. 
 
```{r}
#| eval: true
#| echo: true 
#| message: false
#| code-line-numbers: false
#| classes: code-125

dat <- rio::import("http://people.hsc.edu/faculty-staff/blins/classes/spring17/math222/data/majors.csv")


reg <- lm(`Review Rating` ~ `Discount Applied`, dat)

dat <- dat %>% 
  select(sex, maj, satm, satv, gpa)


colnames(dat) <- c("Gender", "Major", "SAT_math", "SAT_verbal", "GPA")


rio::export(dat, "/home/fsetti/Dropbox/Work/Github repos/Fabio-Setti/static/data/major_gender_SAT.csv", format = "csv")

```

```{r}
# library(FabioFun)

# labels <- c()

# labels_react <- reactbl_col.desc(names = colnames(dat),
#                                  desc =  labels,
#                                  style = "cursor: pointer; text-decoration: none;",
#                                  info_color = "#7a0b80")

reactable::reactable(dat[1:200,],
                     style = list(fontFamily = "Work Sans, sans-serif", fontSize = ".7rem"),
                     pagination = FALSE, highlight = TRUE, height = 300)
```

:::
::::



## Types of Squares 

Something annoying about ANOVAs is that there are multiple "ways" of calculating sums of squares. If for some reason you end up running ANOVAs a bunch, maybe [look more into it](https://stats.stackexchange.com/questions/20452/how-to-interpret-type-i-type-ii-and-type-iii-anova-and-manova){target='_blank'}. We will use Type III sum of squares and call it a day. 


## References 

<div id="refs"> </div>