---
title: "Lab 7: One-Way ANOVA"
author: "Fabio Setti"
institute: "PSYC 6802 - Introduction to Psychology Statistics"
bibliography: Additional files/references.bib
csl: Additional files/apa.csl
title-slide-attributes:
  data-transition: "zoom"
  data-visibility: "uncounted"
format:
   revealjs:
      footer: "PSYC 6802 - Lab 7: One-Way ANOVA"
      width: 1280
      height: 720
      chalkboard: true
      slide-number: c/t 
      theme: Fabio_theme/Fabio_theme.scss
      navigation-mode: linear
      controls: false
      auto-stretch: false
      header-includes:
        - <script src="Fabio_theme/Fabio_theme.js"></script>

editor: source
build: local
---

## Today's Packages and Data 🤗

:::: {.columns}
::: {.column width = '50%'}

```{r}
#| code-fold: true
#| eval: false
#| echo: true
#| code-line-numbers: false
#| code-summary: "Install Packages Code"
#| classes: code-125


# run for packages that you have not installed yet
# install.packages("tidyverse")
install.packages("psych")
install.packages("carData")
```

```{r}
#| eval: true
#| echo: true
#| code-line-numbers: false
#| warning: false
#| classes: code-150

library(psych)
library(carData)
library(tidyverse)
# I am setting a theme for all the plots
theme_set(theme_classic(base_size = 16, accent = "#7a0b80",
                        base_family = 'serif'))
```

</br>

::: {.panel-tabset}

### `psych`

The `psych` package [@Revelle_2024]  includes MANY functions that help with data exploration and running analyses typically done in the social sciences. It is particluarly useful for exploratory factor analyses. 

### `carData`

The `carData` package [@Fox_etal_2022] contains many datasets. Most of these datasets are used in the book titled *an R Companion to Applied Regression*.

:::

:::

::: {.column width = '50%'}
<center style="padding-bottom: 21px;"> [Data]{.data-title} </center>

Today we will work with the `prestige` dataset. This dataset contains the perceived prestige and other variables related to 102 professions in Canada in 1971.
 
```{r}
#| eval: true
#| echo: true 
#| message: false
#| code-line-numbers: false
#| classes: code-125

dat <- rio::import("https://fabio-setti.netlify.app/data/NYC_SAT_data.csv")
```

```{r}
library(labelled)
library(FabioFun)

dat_labs <- rio::import("https://fabio-setti.netlify.app/data/prestige.sav")
labels <- as.character(var_label(dat_labs))

labels <- labels[-1]


labels_react <- reactbl_col.desc(names = colnames(dat),
                                 desc =  labels,
                                 style = "cursor: pointer; text-decoration: none;",
                                 info_color = "#7a0b80")

reactable::reactable(dat,
                     style = list(fontFamily = "Work Sans, sans-serif", fontSize = ".9rem"),
                     pagination = FALSE, highlight = TRUE, height = 300, columns = labels_react)
```

:::
::::




## References 

<div id="refs"> </div>