---
title: 'Lab 12 Activity'
format:
 pdf:
  mainfont: Times New Roman
  sansfont: Times New Roman
  monofont: Courier New
  mathfont: Times New Roman
header-includes:
  - \usepackage{setspace, tikz}
  - \onehalfspacing
editor: source
---



For this lab activity, you will be using the `affairs` data from the `wooldridge` package. Run the following code to install the package and load the data as `dat`.

```{r}
#| warning: false
#| eval: false

install.packages("wooldridge")
dat <- wooldridge::affairs[1:10]
```

The data contains the results of a survey given to 601 married individuals in 1978. Here is a description of the variables in the data:


```{r}
#| echo: false
#| eval: true
#| warning: false

library(tidyverse)
dat <- wooldridge::affairs[1:10]

info <- tibble(Variable = names(dat),
               Description = c("Identifier", 
                               "1 if male", 
                               "Age in years",
                               "years married",
                               "1 if have kids",
                               "5 = very relig., 4 = somewhat, 3 = slightly, 2 = not at all, 1 = anti", 
                               "years schooling",
                               "occupation, reverse Hollingshead scale. See https://fcon_1000.projects.nitrc.org/indi/enhanced/assessments/ses-child.html", 
                               "5 = very hap marr, 4 = happier than avg, 3 = avg, 2 = smewht unhap, 1 = vry unhap",
                               "number of affairs"))
library(kableExtra)

kable(info) %>% 
  column_spec(column = 2, width = "5in") %>% 
  column_spec(column = 1, bold = T) %>% 
  row_spec(0, bold = T)
```

\newpage

**1.** Run a moderation analysis where the dependent variable is `naffairs`, the number of affairs someone has, the focal predictor is `yrsmarr`, numbers of years of marriage, and the moderator is `ratemarr`, individual self report or marriage quality. 

```{r}
#| echo: false
#| eval: false

reg_int <- lm(naffairs ~ yrsmarr*ratemarr, dat)
summary(reg_int)
```

- Is the interaction effect significantly different from 0? What is the interpretation of regression coefficient associated with the interaction effect given that `yrsmarr` is the focal predictor and `ratemarr` is the moderator?

```{r}
#| echo: false
#| eval: false

# Yes, it is significanct. The slope associated with the interaction effect -.04 is the expected change in the slope of yrsmarr per unit change in ratemarr
```


**2.** Wtire out the full regression equation and calculate some simple slopes. 


```{r}
#| echo: false
#| eval: false

# Y = 2.10588 + 0.25029*yrsmarr -0.34164*ratemarr -0.04414*yrsmarr*ratemarr
```

- What is the expected slope of `yrsmarr` when `ratemarr` equals 3?

```{r}
#| echo: false
#| eval: false

coef(reg_int)[2] + coef(reg_int)[4]*3
```

- What is the expected slope of `yrsmarr` when `ratemarr` equals 1?

```{r}
#| echo: false
#| eval: false

coef(reg_int)[2] + coef(reg_int)[4]*1
```


- What is the expected slope of `yrsmarr` when `ratemarr` is at its mean?

```{r}
#| echo: false
#| eval: false

coef(reg_int)[2] + coef(reg_int)[4]*mean(dat$yrsmarr)
```


**3.** Plot the simple slopes of `yrsmarr` at the mean and 1 SD above/below the mean of `ratemarr`. What happens to the relation between `yrsmarr` and `naffairs` as `ratemarr` changes?

```{r}
#| echo: false
#| eval: false

interactions::interact_plot(reg_int, pred = "yrsmarr", modx = "ratemarr")

# as ratemarr decreases the predicted number of affairs per yrsmarr increases
```


**4.** Use `lavaan` to estimate the following mediation model. Make sure to calculate the indirect and total effect.

\begin{tikzpicture}[scale=2.5]
    \node (X) at (0,0) [draw, rectangle] {relig};
    \node (M) at (2,.8) [draw, rectangle] {yrsmarr};
    \node (Y) at (4,0) [draw, rectangle] {ratemarr};
    
    % Arrows
    \draw[-stealth] (X) -- (M) node[midway, above] {a};
    \draw[-stealth] (M) -- (Y) node[midway, above] {b};
    \draw[-stealth] (X) -- (Y) node[midway, above] {c'};
  
\end{tikzpicture}



```{r}
#| echo: false
#| eval: false

lv_med <-"yrsmarr ~ a*relig
           ratemarr ~ c_p*relig + b*yrsmarr
           indirect := a*b
           total := c_p + a*b"

lav_mod_res <- lavaan::sem(lv_med, dat)
```

- Does `yrsmarr` mediate the relation between `relig` and `ratemarr`? Make sure to calculate the appropriate confidence interval


```{r}
#| echo: false
#| eval: false

FabioFun::lavaan_summary(lav_mod_res)

semTools::monteCarloCI(lav_mod_res)

# yes, it does because the confidence intervale of the indirect effect does not include 0.
```

**5.** Let's say that one of your colleagues is not happy with the previous model and says that you should *control* for education, `educ`. To do so, they tell you to run the following model:



\begin{tikzpicture}[scale=2.5]
    % Nodes
    \node (X) at (0,0) [draw, rectangle] {relig};
    \node (M) at (2,.8) [draw, rectangle] {yrsmarr};
    \node (Y) at (4,0) [draw, rectangle] {ratemarr};
    \node (C) at (2,-0.8) [draw, rectangle] {educ};

    % Arrows
    \draw[-stealth] (X) -- (M) node[midway, above] {a};
    \draw[-stealth] (M) -- (Y) node[midway, above] {b};
    \draw[-stealth] (X) -- (Y) node[midway, above] {c'};
    
    % Control variable paths
    \draw[-stealth] (C) -- (M);
    \draw[-stealth] (C) -- (Y);
\end{tikzpicture}



Estimate this model in `lavaan`. Is the indirect effect still significant after controlling for education?

```{r}
#| echo: false
#| eval: false

lv_med <-"yrsmarr ~ a*relig + educ
           ratemarr ~ c_p*relig + b*yrsmarr + educ
           indirect := a*b
           total := c_p + a*b"

# yes         
lav_mod_res <- lavaan::sem(lv_med, dat)
FabioFun::lavaan_summary(lav_mod_res)
```