---
title: "Lab 4 Activity"
format: pdf
editor: source
header-includes:
  - \usepackage{setspace}
  - \onehalfspacing
---

```{r}
#| echo: false
#| eval: true
#| warning: false

library(tidyverse)
```


**1.** Load in the `CO2` data from the `datasets` package by running the following code:

```{r}
#| eval: false
data(CO2, package = "datasets")
help(CO2)
```

Look at the documentation to understand the variables in this dataset. Identify a way to analyze these data using a $t$ test (several possible choices, but `conc` probably won't be an interesting DV). Evaluate each assumption, conduct the $t$ test, compute an effect size, and interpret the results.

```{r}
#| echo: false
#| eval: false

CO2 %>% summary
# Type or Treatment are appropriate IVs for an ind. samples t-test
# uptake is the most obvious DV

## Type as the IV: 

# test for outliers:
CO2 %>% group_by(Type) %>%
  summarize(mean = mean(uptake),
            tmean = mean(uptake, trim = .05), 
            .groups = "keep") # some differences for Quebec
ggplot(CO2, aes(Type, uptake)) + geom_boxplot() 
# possibly some outliers for Quebec

# test for normality:
library(e1071)
CO2 %>% group_by(Type) %>%
  summarize(mean = mean(uptake), sd = sd(uptake),
            skew = skewness(uptake), kurt = kurtosis(uptake),
            n = n(), .groups = "keep")
# some potentially meaningful skew for Quebec, kurt for Mississippi
ggplot(CO2, aes(sample = uptake)) + stat_qq() + 
  stat_qq_line() + facet_wrap(~Type) # some major violations of normality
# but 42/group might mitigate against this violation

# test for homogeneity of variance:
library(car)
leveneTest(uptake ~ Type, CO2)
# homogeneity not violated, but we'll use Welch anyways

# t-test (two-tailed, Welch)
t.test(uptake ~ Type, CO2)

# effect size - put pooled standard deviation in the denominator
CO2 %>% group_by(Type) %>% summarize(var(uptake), n(), .groups = "keep")
sp <- sqrt((41 * 93.6 + 41 * 61.1) / (42 + 42 - 2))
(33.54286 - 20.88333) / sp

# mean CO2 uptake rate in Quebec is significantly higher than in Mississippi with a very large effect size (d = 1.44)


## Treatment as the IV:

# test for outliers:
CO2 %>% group_by(Treatment) %>%
  summarize(mean = mean(uptake),
            tmean = mean(uptake, trim = .05), 
            .groups = "keep") # no major differences
ggplot(CO2, aes(Treatment, uptake)) + geom_boxplot() 
# no outliers shown

# test for normality:
library(e1071)
CO2 %>% group_by(Treatment) %>%
  summarize(mean = mean(uptake), sd = sd(uptake),
            skew = skewness(uptake), kurt = kurtosis(uptake),
            n = n(), .groups = "keep")
# some potentially meaningful skew and kurt for both groups
ggplot(CO2, aes(sample = uptake)) + stat_qq() + 
  stat_qq_line() + facet_wrap(~Treatment) # some major violations of normality
# but 42/group might mitigate against this violation

# test for homogeneity of variance:
library(car)
leveneTest(uptake ~ Treatment, CO2)
# homogeneity not violated, but we'll use Welch anyways

# t-test (two-tailed, Welch)
t.test(uptake ~ Treatment, CO2)

# effect size - put pooled standard deviation in the denominator
CO2 %>% group_by(Treatment) %>% summarize(var(uptake), n(), .groups = "keep")
sp <- sqrt((41 * 94.2 + 41 * 118) / (42 + 42 - 2))
(30.64286 - 23.78333) / sp

# mean CO2 uptake rate for nonchilled treatment is significantly higher than in for the chilled treatment with a large effect size (d = .67)
```

**2.** Load in the `sleep` data from the `datasets` package by running the following code:

```{r}

data(sleep, package = "datasets")

# This makes analyzing the data easier

sleep <- sleep %>%  
            tidyr::pivot_wider(id_cols = ID, 
                               names_from = group, 
                               names_prefix = "group",
                               values_from = extra)
```

Look at the documentation to understand the variables in this dataset (but try not to peek at the examples section of the documentation). Identify a way to analyze these data using a $t$ test. Evaluate each assumption, conduct the $t$ test, compute an effect size, and interpret the results.


```{r}
#| echo: false
#| eval: false
data(sleep, package = "datasets")
sleep <- sleep %>% 
  pivot_wider(id_cols = ID, 
              names_from = group, 
              names_prefix = "group",
              values_from = extra)
sleep %>% summary

# extra should be the DV
# group is an appropriate IV for paired samples t-test (see documentation)

# want to evaluate the normality of difference scores
sleep <- sleep %>% mutate(diff = group2 - group1)

# trimmed mean may not be appropriate because N is so small
ggplot(sleep, aes(diff)) + geom_boxplot()
# possibly likely outliers, but hard to draw any definitive conclusions

ggplot(sleep, aes(sample = diff)) + stat_qq() + stat_qq_line() # probably non-normal

# an upper-tailed test may be appropriate if we are only concerned about gains
t.test(sleep$diff, alternative = "greater")

# effect size
4.0621 / sqrt(10)

# there is a significant increase in sleep with drug 2 over drug 1 with a large effect size (d = 1.28), but sample size is small, so we are cautious about this conclusion
```


**3.** The `t.test()` function also includes an `alternative =` argument. What does that argument do? How do your results differ if you try different values for the `alternative =` argument? Make sure to read the help menu of the `t.test()` function.

```{r}
#| echo: false
#| eval: false

# the alternative =  argument allows to specified directional hypotheses. The p-value of directional hypotheses will always be twice as small as the p-value of non directional hypotheses. 
```


**4.** For (1) and (2), use `ggplot` to create interesting graphs to illustrate your results (what might you include in a thesis or published paper?).

```{r}
#| echo: false
#| eval: false

## depends on what students chose for 1 and 2, many possible answers
```