---
title: "Lab 3: Two-Predictor Regression"
author: "Fabio Setti"
institute: "Spring 2026"
subtitle: "PSYC 7804 - Regression with Lab"
bibliography: Additional files/R packages.bib
csl: Additional files/apa.csl
notice: |
  @Becker_etal_2024a
  @Wickham_RStudio_2023
title-slide-attributes:
  data-transition: "zoom"
  data-visibility: "uncounted"
format:
   revealjs:
      footer: "PSYC 7804 - Lab 3: Two-Predictor Regression"
      width: 1280
      height: 720
      chalkboard: true
      slide-number: c/t 
      theme: Fabio_theme/Fabio_theme.scss
      navigation-mode: linear
      controls: false
      auto-stretch: false
      header-includes:
        - <script src="Fabio_theme/Fabio_theme.js"></script>

editor: source
---
## Today's Packages and Data 🤗


:::: {.columns}
::: {.column width="50%"}


```{r}
#| code-fold: true
#| eval: false
#| echo: true
#| code-line-numbers: false
#| code-summary: "Install Packages Code"
#| classes: code-150

install.packages("GGally")
install.packages("devtools")
install.packages("flextable")
# requires devtools
devtools::install_github("quinix45/FabioFun")

```


```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

library(tidyverse)
library(GGally)
library(FabioFun)
theme_set(theme_classic(base_size = 16, accent = "#3492eb", base_family = 'serif'))
```


</br>

<div style="font-size: 26px">

::: {.panel-tabset}

### `GGally`

The `GGally` package [@Schloerke_etal_2024] builds upon `ggplot2` and includes many functions for creating complex plots. 

### `devtools`

The `devtools` package [@Wickham_etal_2022] helps developing R packages. It is generally used to download R packages from [github](https://github.com){target='_blank'}.

### `FabioFun`

This is my personal package where I save functions that I use for research/teaching/convenience. You can run `devtools::install_github("quinix45/FabioFun")` to install it. 

### `flextable`

The `flextable` package [@Gohel_etal_2024] allows to create highly customizable tables. The function that I use to create APA style Word tables makes heavy use of this package.

:::

</div>

:::


::: {.column width="50%"}

<center style="padding-bottom: 21px;"> [Data]{.data-title} </center>


Today we'll be looking at data from the [2024 World Happiness report](https://www.worldhappiness.report/ed/2024/){target='_blank'}. 
 
```{r}
#| eval: true
#| echo: true 
#| message: false
#| code-line-numbers: false
#| classes: code-125

dat <- rio::import("https://fabio-setti.netlify.app/data/World_happiness_2024.csv")
```

```{r}
library(FabioFun)


labels <- c("Country name",
            "Region of the World",
            "Country happiness score. Corresponds to the sum of the other variables",
            "Log of country Gross Domestic Product (GDP)",
            "Perceived Social Support in the Country",
            "Country Healthy life Expectancy",
            "Perceived Freedom in the Country",
            "Perceived Generosity in the Country",
            "Perceived Corruption of the Country")

labels_react <- reactbl_col.desc(names = colnames(dat),
                                 desc =  labels,
                                 style = "cursor: pointer; text-decoration: none;",
                                 info_color = "#7a0b80")

reactable::reactable(dat,
                     defaultColDef = colDef(minWidth = 170),
                     style = list(fontFamily = "Work Sans, sans-serif", fontSize = "1rem"),
                     pagination = FALSE, highlight = TRUE, height = 300, columns = labels_react)
```

:::

::::


## Predicting Country Happiness

In our data we have many variables that are likely related to happiness. I will just pick 2 of them for the purpose of this Lab. Let's say that we want to evaluate how `Freedom` and `Corruption` predict `Happiness`. If we run two separate regressions, we find that both have reasonably high $R^2$ values:

:::: {.columns}
::: {.column width = '50%'}

::: {.fragment fragment-index=1}


```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-115

reg_free <- lm(Happiness_score ~ Freedom, dat)
summary(reg_free)
```

:::

:::
::: {.column width = '50%'}

::: {.fragment fragment-index=2}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-115

reg_corr <- lm(Happiness_score ~ Corruption, dat)
summary(reg_corr)
```

:::

:::
::::


## Well, That's just the correlation 🤷

As we have learned, regression with one predictor is just a contrived way of calculating correlations among variables. We can better visualize the relation among out 3 variables with the `ggpairs()` function from `GGally`:

::: {.fragment fragment-index=1}

:::: {.columns}
::: {.column width = '35%'}

```{r}
#| eval: false
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

# here I am selecting the 3 columns that I wantr to plot
GGally::ggpairs(dat[,c(3, 7,9)])
```

- **Lower part:** scatterplot between variables
- **Diagonal:** variable distribution
- **Upper part:** correlation between variables

Notably, `Freedom` and `Corruption` are also correlated.

:::
::: {.column width = '65%'}

```{r}
#| eval: true
#| echo: false
#| code-line-numbers: false
#| classes: code-125

ggpairs(dat[,c(3, 7,9)])
```

:::
::::

:::

::: {.fragment fragment-index=2}

Separate regressions miss the points that our predictors share information! 

:::

## Multiple Independent Variables

:::: {.columns}
::: {.column width = '52%'}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

reg_multi <- lm(Happiness_score ~ Freedom + Corruption, dat)
summary(reg_multi)
```

:::
::: {.column width = '48%'}

We can include both predictors in our regression by adding variables in the `lm()` function like so `lm(DV ~ IV1 + IV2 + ..., dat)`. 


::: {.fragment fragment-index=1}
Note that both slopes have changed and are not as large as they were in the individual regressions. The regression equations is now:

 <p style='font-size: 20px'>   $$\mathrm{Happiness} = 5.09 + 4.03 \times \mathrm{Freedom} - 2.44 \times \mathrm{Corruption}$$ </p>

:::

::: {.fragment fragment-index=2}
What is going on here? 🤔 a graphical illustration may provide some intuition!
:::

:::
::::


## Individual regression plots

Below I plot the individual regression lines for `Freedom` and `Corruption` predicting `Happiness`. 

:::: {.columns}
::: {.column width = '50%'}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot Code"

ggplot(dat,  
 aes(x = Freedom, y = Happiness_score)) +
 geom_point() + 
    geom_smooth(method = "lm",  
                se = FALSE)
```

:::
::: {.column width = '50%'}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot Code"

ggplot(dat,  
 aes(x = Corruption, y = Happiness_score)) +
 geom_point() + 
    geom_smooth(method = "lm",  
                se = FALSE)
```



:::
::::

::: {.fragment fragment-index=1}

Now, what happens if we try to overlay the regression lines that we got from our multiple regression? Will they "fit the data" better that the regression lines shown here?

:::



## Adding Multiple regression Lines

if we add our multiple regression lines to the plot... (🥁 *drum roll*)

::: {.fragment fragment-index=1}

:::: {.columns}
::: {.column width = '50%'}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot Code"

ggplot(dat,  
 aes(x = Freedom, y = Happiness_score)) +
 geom_point() + 
    geom_smooth(method = "lm",  
                se = FALSE) +
    geom_abline(intercept = coef(reg_multi)[1], 
                slope = coef(reg_multi)[2])

```

:::
::: {.column width = '50%'}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot Code"

ggplot(dat,  
 aes(x = Corruption, y = Happiness_score)) +
 geom_point() + 
    geom_smooth(method = "lm",  
                se = FALSE)+
    geom_abline(intercept = coef(reg_multi)[1], 
                slope = coef(reg_multi)[3])

```

:::
::::

:::

::: {.fragment fragment-index=2}

Wait, we are way off now 🤨 And here I thought multiple regression would be better 😑 ...Unsurprisingly, the multiple regression is fine, our *perspective* is the problem!

:::

## A better Perspective

We have *3* variables, `Freedom`, `Corruption`, and `Happiness`, but we are looking at them in *2* dimensions with the individual scatterplots. We need a **3D** scatterplot!



:::: {.columns}
::: {.column width = '40%'}

::: {.fragment fragment-index=1}

```{r}
#| eval: false
#| echo: true 
#| code-line-numbers: false
#| classes: code-125


# run "devtools::install_github("quinix45/FabioFun")" if you haven't installed the package yet
library(FabioFun)

nice_3D_plot(y = dat$Happiness_score,
             x1 = dat$Freedom,
             x2 = dat$Corruption,
             dot_labels = dat$Country_name,
             axis_names = c("Happiness", 
                            "Freedom", 
                            "Corruption"))
```

What is happening is that the scatterplots are *only the sides* of this box ‼️

:::

::: {.fragment fragment-index=2}
Multiple regression works on this plot on the right, not the individual ones; multiple regression *operates on a different dimension*.
:::

:::
::: {.column width = '60%'}

::: {.fragment fragment-index=1}

```{r out.width = "100%"}

library(FabioFun)


nice_3D_plot(y = dat$Happiness_score,
             x1 = dat$Freedom,
             x2 = dat$Corruption,
             dot_labels = dat$Country_name,
             axis_names = c("Happiness", 
                            "Freedom", 
                            "Corruption")) %>%  
                    bslib::card(full_screen = TRUE)
```

:::

:::
::::

## The Regression *Plane*

The results from multiple regression create a **regression plane**, which is the 3D version of a regression line. 

:::: {.columns}
::: {.column width = '40%'}

::: {.fragment fragment-index=1}

```{r}
#| eval: false
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

nice_3D_plot(y = dat$Happiness_score,
             x1 = dat$Freedom,
             x2 = dat$Corruption,
             dot_labels = dat$Country_name,
             axis_names = c("Happiness", 
                            "Freedom", 
                            "Corruption"),
             reg_plane = TRUE)
```


The plane works exactly as the regression line. its surface is the best predition of `happiness` given a value of `Freedom` and `Corruption`. The residuals are now the distance from any of the dots to the plane! 

:::


::: {.fragment fragment-index=2}
**Question:** What would the plot look like with one extra variable?
:::

:::

::: {.column width = '60%'}

::: {.fragment fragment-index=1}

```{r out.width = "100%"}

nice_3D_plot(y = dat$Happiness_score,
             x1 = dat$Freedom,
             x2 = dat$Corruption,
             dot_labels = dat$Country_name,
             axis_names = c("Happiness", 
                            "Freedom", 
                            "Corruption"),
             reg_plane = TRUE) %>%  
                    bslib::card(full_screen = TRUE)
```

:::

:::
::::


## Interpreting Multiple Regression

:::: {.columns}
::: {.column width = '50%'}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

summary(reg_multi)
```


:::
::: {.column width = '50%'}

Now that we have looked at our results graphically for a while, let's interpret the numbers that we get.

::: {.fragment fragment-index=1}

- **intercept**: the expected value of `Happiness` when both `Freedom` and `Corruption` are at 0.


- **Slope of `Freedom`**: the expected change in `Happiness` for 1-unit increase in `Freedom`, controlling for `Corruption`. 

- **Slope of `corruption`**: the expected change in `Happiness` for 1-unit increase in `Corruption`, controlling for `Freedom`.

:::

:::
::::

::: {.fragment fragment-index=5}
**Note:** Instead of "controlling for", you may also hear "accounting for" or "holding constant". They are all the same thing. 
:::

## Variance as what we cannot predict

At this point, a quick detour about variance is in order. In Psychology we work with **random variables**. The word <u>*random*</u> is important because it means that there is something that we cannot predict about the variable (e.g., we cannot perfectly predict depression, nor will we ever be able to).

:::: {.columns}
::: {.column width = '40%'}

::: {.fragment fragment-index=1}
However, given a sample, we can quantify the unpredictability of a variable. The **variance** (or equivalently the SD) measure how uncertain a variable is!
:::

::: {.fragment fragment-index=2}

We can see that the red distribution has a larger variance than the blue distribution. That is, the red distributrion has more unpredictability.

:::

::: {.fragment fragment-index=3}

(by the way, generally [Entropy](https://en.wikipedia.org/wiki/Entropy_(information_theory)){target='_blank'} is used to quantify unpredictability of a random variable)

:::

:::
::: {.column width = '60%'}

::: {.fragment fragment-index=2}

```{r}
#| eval: true
#| echo: true 
#| code-fold: true
#| code-summary: Plot code
 
ggplot() +
 geom_function(fun = dnorm, 
               args = list(mean = -4,
                           sd = 1), color = "blue") + 
 geom_function(fun = dnorm, 
               args = list(mean = 4,
                           sd = 2), color = "red") +
                        theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.line.y =  element_blank()) + xlim(c(-10, 10))
```

:::

:::
::::


## R<sup>2</sup> as Uncertainty Explained

After running a regression we can calculate a statistic called $R^2$, which quantifies the predictive power of the regression. $R^2$ is also often termed "variance explained". Let's figure out why:

:::: {.columns}
::: {.column width = '70%'}

::: {.fragment fragment-index=1}

Our regression where $Y =$ `Happiness` had an $R^2$ of:

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

summary(reg_multi)$r.squared
```

:::

::: {.fragment fragment-index=2}

Now, the variance of `Happiness_score` originally was:

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

var(dat$Happiness_score)
```

:::

::: {.fragment fragment-index=3}

Additionally, we can think of the **variance of the residuals** as quantifying what we could not predict:

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

var(reg_multi$resid)
```

:::

:::
::: {.column width = '30%'}

::: {.fragment fragment-index=4}

So, we can see that the variance of our $Y$ variable, `Happiness`, has been reduced by 

$$ 1 - \frac{0.73}{1.395} = 0.475\%  = R^2 $$

:::


::: {.fragment fragment-index=5}

$R^2$ is an effect size and it tells us how much uncertainty in the $Y$ variable has been reduced by the predictors.

:::

:::
::::


## A graphical look at  R<sup>2</sup> 


:::: {.columns}
::: {.column width = '30%'}

We can plot the density of `Happiness` and the residuals just like we plotted the two distributions a few slides ago.


::: {.fragment fragment-index=1}

Just as before, you can see that the two distributions vidibly differ in variance. The residuals (blue density) have less variance, $47.5\%$ less to be precise.

:::

::: {.fragment fragment-index=2}

**Question:** What would it mean if the residuals had no variance at all? 🤔

:::

:::
::: {.column width = '70%'}


```{r}
#| eval: true
#| echo: true 
#| code-fold: true
#| code-summary: Plot code
 
ggplot(dat) +
  geom_density(
    aes(x = Happiness_score, color = "Happiness (Original Variable)")
  ) +
  geom_density(
    aes(x = reg_multi$resid, color = "Regression residuals")
  ) +
  scale_color_manual(
    name = "",
    values = c(
      "Happiness (Original Variable)" = "red",
      "Regression residuals" = "blue"
    )
  ) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y  = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y  = element_blank(),
    legend.position = "bottom"
  ) +
  xlim(c(-5, 10))

```

:::
::::

## Significance Test for R<sup>2</sup> 

We have seen that we can individually test whether each of the regression coefficients are significantly different from 0 with *t*-tests. However, this may become a bit cumbersome when we introduce many predictors in our regression.

:::: {.columns}
::: {.column width = '50%'}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

summary(reg_multi)$coef
```


:::
::: {.column width = '50%'}


::: {.fragment fragment-index=1}

If you are not as interested to every individual variable, another approach is to test whether the regression $R^2$ is significantly different from 0.

:::


:::
::::

::: {.fragment fragment-index=2}

:::: {.columns}
::: {.column width = '50%'}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

summary(reg_multi)$r.squared

summary(reg_multi)$fstatistic

pf(62.05882, 2, 137, lower.tail = FALSE)
```


:::
::: {.column width = '50%'}

The significance test for the regression $R^2$ is also found at the end of the `summary()` output. Here we have that $R^2 = .48, F(2, 137) = 62.06, p < .001$.

This means that our regression model explains a statistically significant proportion of variance in the outcome, `Happiness`.

:::
::::

:::

## APA Style Table for Regression Results

because I hate creating tables manually, at some point I made function that does that for me ✨ Also, creating big tables by hand is generally not a good idea: you waste time and likely make typos.

:::: {.columns}
::: {.column width = '70%'}

::: {.fragment fragment-index=1}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

reg_tab <- FabioFun::regression_flextable(reg_multi, 
                                var_names = c("Freedom", "Corruption"), 
                                intercept = TRUE)
reg_tab
```

:::

::: {.fragment fragment-index=2}

```{r}
#| eval: false
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

# save to Word document (will be saved to current working directory)
flextable::save_as_docx(reg_tab, path = "table.docx")
```


Feel free to use this instead of creating regression tables manually.

:::

:::
::: {.column width = '30%'}

::: {.fragment fragment-index=1}
the `regression_flextable()` function I made takes in an `lm()` object and has some optional arguments (see help page). Here I rename the regression coefficients with the `var_names =` argument.

:::


::: {.fragment fragment-index=2}
You can save the table as a `.docx` file (Word document) by using the `save_as_docx()` function from the `flextable()` package. You specify the file name in the `path =` argument.
:::

:::
::::

## References 

<div id="refs"> </div>


# Appendix: `nice_3D_plot()` Showcase


## Example with World Happiness Data

You can also add groups to the dots of the 3D plot. Here I am using `Region`.

:::: {.columns}
::: {.column width="50%"}

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: false
#| classes: code-125

# selecting just the variables of interest makes code cleaner later
reg_vars <- dat[, c("Happiness_score", "Log_GDP", "Freedom")]

nice_3D_plot(y = reg_vars$Happiness_score,
             x1 = reg_vars$Log_GDP,
             x2 = reg_vars$Freedom,
             dot_labels = dat$Country_name,
             # add groups
             groups = dat$Region,
             axis_names = c("Happiness", 
                            "GDP", 
                            "Freedom"),
             plane_res = 20,
             reg_plane = FALSE)

```


There are a bit too many regions, so this is not very practical in this case, but this just an example.

:::


::: {.column width="50%"}

```{r}
#| eval: true
#| echo: false
#| code-line-numbers: false
#| classes: code-125

reg_vars <- dat[, c("Happiness_score", "Log_GDP", "Freedom")]

nice_3D_plot(y = reg_vars$Happiness_score,
                            x1 = reg_vars$Log_GDP,
                            x2 = reg_vars$Freedom,
                            dot_labels = dat$Country_name,
                            groups = dat$Region,
                              axis_names = c("Happiness", 
                                             "GDP", 
                                             "Freedom"),
                              plane_res = 20,
                              reg_plane = FALSE) %>%  
                    bslib::card(full_screen = TRUE)

```

:::
::::

## Example with iris Data

A more practical example with the [iris](https://en.wikipedia.org/wiki/Iris_flower_data_set){target="_blank"} dataset, Let's start with just providing 3 variables.

:::: {.columns}
::: {.column width="40%"}

```{r}
#| eval: false
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

nice_3D_plot(y = iris$Sepal.Length,
             x1 = iris$Petal.Length,
             x2 = iris$Petal.Width)
```



:::
::: {.column width="60%"}


```{r}
#| eval: true
#| echo: false
#| code-line-numbers: false
#| classes: code-125

nice_3D_plot(y = iris$Sepal.Length,
             x1 = iris$Petal.Length,
             x2 = iris$Petal.Width) %>%  
  bslib::card(full_screen = TRUE)
```

:::
::::


## Name Axes

Let's name our axes:

:::: {.columns}
::: {.column width="40%"}

```{r}
#| eval: false
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

nice_3D_plot(y = iris$Sepal.Length,
             x1 = iris$Petal.Length,
             x2 = iris$Petal.Width,
             axis_names = c("SL", "PL", "PW"))
```



:::
::: {.column width="60%"}


```{r}
#| eval: true
#| echo: false
#| code-line-numbers: false
#| classes: code-125

nice_3D_plot(y = iris$Sepal.Length,
             x1 = iris$Petal.Length,
             x2 = iris$Petal.Width,
             axis_names = c("SL", "PL", "PW")) %>%  
  bslib::card(full_screen = TRUE)
```

:::
::::




## Add Groups 

Let's add colors for the 3 different species of flowers:

:::: {.columns}
::: {.column width="40%"}

```{r}
#| eval: false
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

nice_3D_plot(y = iris$Sepal.Length,
             x1 = iris$Petal.Length,
             x2 = iris$Petal.Width,
             axis_names = c("SL", "PL", "PW"),
             groups = iris$Species)
```


:::
::: {.column width="60%"}


```{r}
#| eval: true
#| echo: false
#| code-line-numbers: false
#| classes: code-125

nice_3D_plot(y = iris$Sepal.Length,
             x1 = iris$Petal.Length,
             x2 = iris$Petal.Width,
             groups = iris$Species,
             axis_names = c("SL", "PL", "PW")) %>%  
  bslib::card(full_screen = TRUE)
```

:::
::::

## Add Regression Plane

Let's add a regression plane:

:::: {.columns}
::: {.column width="40%"}

```{r}
#| eval: false
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

nice_3D_plot(y = iris$Sepal.Length,
             x1 = iris$Petal.Length,
             x2 = iris$Petal.Width,
             axis_names = c("SL", "PL", "PW"),
             groups = iris$Species,
             reg_plane = TRUE)
```


:::
::: {.column width="60%"}


```{r}
#| eval: true
#| echo: false
#| code-line-numbers: false
#| classes: code-125

nice_3D_plot(y = iris$Sepal.Length,
             x1 = iris$Petal.Length,
             x2 = iris$Petal.Width,
             groups = iris$Species,
             axis_names = c("SL", "PL", "PW"),
             reg_plane = TRUE)%>%  
  bslib::card(full_screen = TRUE)
```

:::
::::


## Pass Arguments to `plot_ly()`

You can also pass arguments to the `plot_ly()` function. Let's make the dots more transparent by passing the `opacity =` argument to `plot_ly()`:

:::: {.columns}
::: {.column width="40%"}

```{r}
#| eval: false
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

nice_3D_plot(y = iris$Sepal.Length,
             x1 = iris$Petal.Length,
             x2 = iris$Petal.Width,
             axis_names = c("SL", "PL", "PW"),
             groups = iris$Species,
             reg_plane = TRUE,
             opacity = .3)
```


See [here](https://www.r-bloggers.com/2015/02/r-three-dots-ellipsis/){target="_blank"} for what "passing arguments" to a function within a function means in R. 


:::
::: {.column width="60%"}


```{r}
#| eval: true
#| echo: false
#| code-line-numbers: false
#| classes: code-125


nice_3D_plot(y = iris$Sepal.Length,
             x1 = iris$Petal.Length,
             x2 = iris$Petal.Width,
             axis_names = c("SL", "PL", "PW"),
             groups = iris$Species,
             reg_plane = TRUE,
             opacity = .3)%>%  
  bslib::card(full_screen = TRUE)
```

:::
::::


## Add Interaction Term to Regression Plane

You can also specify an interaction between the two predictors and get an interaction plane. You will see that the reression plane now bends:

:::: {.columns}
::: {.column width="40%"}

```{r}
#| eval: false
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

nice_3D_plot(y = iris$Sepal.Length,
             x1 = iris$Petal.Length,
             x2 = iris$Petal.Width,
             axis_names = c("SL", "PL", "PW"),
             groups = iris$Species,
             reg_plane = TRUE,
             opacity = .3,
             interaction = TRUE)
```

See [Lab 9](https://raw.githack.com/quinix45/PSYC-7804-Regression-Lab-Slides/main/Slides%20Files/Lab%209.html#/title-slide){target="_blank"} if you are curious to know what is happening. 


:::
::: {.column width="60%"}

```{r}
#| eval: true
#| echo: false
#| code-line-numbers: false
#| classes: code-125


nice_3D_plot(y = iris$Sepal.Length,
             x1 = iris$Petal.Length,
             x2 = iris$Petal.Width,
             axis_names = c("SL", "PL", "PW"),
             groups = iris$Species,
             reg_plane = TRUE,
             opacity = .3,
             interaction = TRUE)%>%  
  bslib::card(full_screen = TRUE)
```

:::
::::
